### 2009-04-06 talk/icooolps2009-dotnet/intro.tex
remove commented out text, a redundant paragraph and an XXX

When using PyPy, the first step is to write an interpreter for the chosen language.  Since it
must be fed to the translation toolchain, the interpreter has to be written in
RPython.  Then, to guide the process, we need to add few manual
annotations to the interpreter, in order to teach the JIT generator which
information is important to know at compile-time.  Annotations are inserted
as \emph{hints}, as described in section \ref{sec:hints}.
From these hints, PyPy will statically generate an interpreter and a JIT
compiler in a single executable (here a .NET executable).

\anto{maybe we can avoid to talk about hints?}

\commentout{
It is important to distinguish between three distinct phases of execution:

\begin{enumerate}
\item \textbf{Translation-time}: when the translation toolchain runs and the
  JIT compiler is automatically generated from the interpreter.  The result is
  an executable that can be used to run programs written in the chosen
  language.
\item \textbf{Compile-time}: when the JIT compiler runs, generating executable
  code on the fly.
\item \textbf{Runtime}: when the code generated at compile-time is executed.
\end{enumerate}

Note that in this schema translation-time happens only once, on the
developer's machine.  By contrast, compile-time and runtime happens every time
the user wants to run some program.
}

\commentout{
By emitting code at runtime, JIT compilers can exploit extra
knowledge compared to traditional static compilers -- more than just the
type of the function arguments.  However, special care is needed to
choose a strategy for JIT compilation that lets the compiler take the
best of this advantage.
}

The interesting property of the generated JIT compiler is to delay the
compilation until it knows all the information needed to generate
efficient code.  In other words, at runtime, when the interpreter notice
that it is useful to compile a given piece of code, it sends it to the
JIT compiler; however, if at some point the JIT compiler does not know
about something it needs, it generates a callback into itself and stops
execution.

@@

Implementing such a behavior requires a tight coupling between compile-time
and run-time: a \emph{callback} is put in the generated code, which can invoke
the compiler again.  When the callback is actually reached at run-time, and
only then, the compiler resumes and uses the knowledge of the actual run-time
value to generate more code.

The new generated code is potentially different for each run-time value seen.
This implies that the generated code needs to contain some sort of updatable
switch, or \emph{flexswitch}, which can pick the right code path based on the
run-time value.  Typically, the value we switch on is the runtime dynamic type
of a value, so that the JIT compiler has all information needed to produce
very good code for that specific case.

\commentout{
The primitive to do this sort of interleaving is called promotion,
it is described in Section \ref{sec:promotion}.

One of the most important optimizations that the generated JIT does is removing
unnecessary allocations. This is described in Section \ref{sec:virtuals}
}

@@

\anto{XXX: mention virtuals/escape analysis?}

>>>>

When using PyPy, the first step is to write an interpreter for the chosen language.  Since it
must be fed to the translation toolchain, the interpreter has to be written in
RPython.  Then, to guide the process, we need to add few manual
annotations to the interpreter, in order to teach the JIT generator which
information is important to know at compile-time.  
From these annotations, PyPy will statically generate an interpreter and a JIT
compiler in a single executable (here a .NET executable).

The interesting property of the generated JIT compiler is to delay the
compilation until it knows all the information needed to generate
efficient code.  In other words, at runtime, when the interpreter notice
that it is useful to compile a given piece of code, it sends it to the
JIT compiler; however, if at some point the JIT compiler does not know
about something it needs, it generates a \emph{callback} into itself and stops
execution.

@@

Potentially, the JIT compiler generates new code for each different run-time
value seen in variables it is interested in.
This implies that the generated code needs to contain some sort of updatable
switch, or \emph{flexswitch}, which can pick the right code path based on the
run-time value.  Typically, the value we switch on is the runtime dynamic type
of a value, so that the JIT compiler has all information needed to produce
very good code for that specific case.

@@
